MSD panels 1, 2 and 3 Explant Supernatant

### 2016-09-22 MSD_panels_1_2_3_explant_sup_analysis.Rmd
changed to doing subtracting log virus-log mock and then getting ratio of 24hr  /3hr  by subtraction because loga - logb = log(a/b)

mergeMock <- mergeMock %>%
  mutate(mockSubtracted = avgEstConcVirus - avgEstConcMock)

@@

## Plots of Analyte Concentrations by Donor, Timepoint and Condition
We did a 1:5 dilution of the samples so I multiplied the concentrations by 5 to get an estimate of the stock supernatent concentration.

@@

#tiles represent foldchange of Mock subtracted 24hr data over mock-subtracted 3hr. So if 24hr>3hr, you get a ratio <1 and if 24hr <3hr you get a ratio >1

@@

timepoint3<- no8mergeMock %>%
  filter(Timepoint == "3")%>%
  select(Donor, Assay, Sample, Virus, mockSubtracted)

#extract 24 hr data and select a few columns
timepoint24<- no8mergeMock %>%
  filter(Timepoint =="24")%>%
  select(Donor, Assay, Sample, Virus, mockSubtracted)

@@

#Fix up colnames      
names(merge24And3)[4:7]<- c("Sample.24", "mockSubtracted.24","Sample.3", "mockSubtracted.3")
```

```{r adjust zeros, check NAs, echo=FALSE}

#Problems with zeros:
#If data from both timepoints is zero we'll get NaN for the ratio

#If data from time 3 = 0, we'll get Inf or -Inf
#If data from time 24 = zero, we'll get zero

#I'm going to change all zeros to 0.001 to avoid these problems #I think this is ok since a heatmap isnt really quantitative

adjZerosMerge24And3<- merge24And3 %>%
  mutate(mockSubtracted.24 = ifelse(mockSubtracted.24 == 0, 0.001,mockSubtracted.24))%>%
  mutate(mockSubtracted.3 = ifelse(mockSubtracted.3 == 0, 0.001,mockSubtracted.3))

@@

#Do the division to get a foldchange. Since we are looking at ratios, it doesnt give any info if the number is negative, plus we can't show that on a log scale. 
#So I'm taking the abs value of the ratio.
#and also making a column for the log10 of the ratio for plotting.

adjZerosMerge24And3 <- adjZerosMerge24And3 %>%
  mutate(AbsVal_mockSub24_Over_mockSub3 = abs(mockSubtracted.24/mockSubtracted.3))%>%
  mutate(log10ratio = log10(AbsVal_mockSub24_Over_mockSub3))

@@

ggplot(adjZerosMerge24And3, aes(x = Donor, y = Assay))+
  geom_tile(aes(fill = log10ratio), color = "white")+
  scale_fill_gradient(low = "green", high = "red")+
  facet_wrap(~Virus)+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")

>>>>

#do the subtraction, with the result being the difference between the log10 of virus - log 10 Mock
mergeMock <- mergeMock %>%
  mutate(logMockSubtracted = log10(avgEstConcVirus) - log10(avgEstConcMock))

@@

We did a 1:5 dilution of the samples so I multiplied the concentrations by 5 to get an estimate of the stock supernatent concentration.

@@

#tiles represent foldchange of Mock subtracted 24hr data over mock-subtracted 3hr. So if 24hr>3hr, you get a ratio <1 and if 24hr <3hr you get a ratio >1.

#recall that log(a)-log(b) = log(a/b) so I'm going to do subtration, not division to get the log ratio FC.

@@

timepoint3<- no8mergeMock %>%
  filter(Timepoint == "3")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

#extract 24 hr data and select a few columns
timepoint24<- no8mergeMock %>%
  filter(Timepoint =="24")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

@@

#Fix up colnames      
names(merge24And3)[4:7]<- c("Sample.24", "logMockSubtracted.24","Sample.3", "logMockSubtracted.3")
```

@@

#Do the subtraction to get a foldchange.

merge24And3 <- merge24And3 %>%
  mutate(logMockSub24_minus_logMockSub3 = logMockSubtracted.24-logMockSubtracted.3)
```

@@

ggplot(merge24And3, aes(x = Donor, y = Assay))+
  geom_tile(aes(fill = logMockSub24_minus_logMockSub3), color = "white")+
  scale_fill_gradient(low = "green", high = "red")+
  facet_wrap(~Virus)+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")
